### 2020-05-29 serial.tex
methods

One source of confusion arises from an apparent discrepancy between the serial-interval and generation-interval viewpoints.
When the epidemic is growing exponentially, the spread of infection can be characterized as a \emph{renewal process} based on previous incidence of infection, the associated generation-interval distribution, and the average infectiousness of an infected individual.
It is well established that this renewal formulation allows us to link the exponential growth rate of an epidemic $r$ with its reproduction number \RR using the generation-interval distribution \citep{wallinga2007generation}.
By definition, the serial-interval distribution describes the renewal process between symptomatic cases based on their symptom onset dates;
since both renewal processes, based on generation- and serial-interval distributions, describe the same underlying system of exponential growth, both should provide the same correct link between \RR and the epidemic growth rate $r$.
In contexts where the distributions are expected to be different, current theory has no explanation for how these differing distributions could provide identical estimates of \RR from $r$. 
Some studies have further suggested that using serial intervals can give different estimates of \RR \citep{britton2019estimation, ganyani2020estimating}.

@@

\eref{backward} suggests that backward delay distributions change over time even if their corresponding forward delay distribution does not change.
Backward delay distributions depend on changes in the primary cohort size over time due to conditionality of observations:
Conditioning on individuals whose secondary events have occurred at the same time means that we tend to observe shorter (or longer) inter-event delays when cohort size has been increasing (decreasing) through time.
When incidence is growing exponentially, we can calculate the amount of bias exactly.
Assuming that the forward delay distribution remains constant during the exponential growth phase ($f_p(\tau) \approx f_0(\tau)$), we can substitute $\psize(t) = \psize(0) \exp(rt)$ in \eref{backward} to obtain:
\begin{align}
b_0(\tau) &= \left[\psize(0)/\ssize(0)\right] \exp(-r\tau) f_0(\tau)
\label{eq:backexp}
\end{align}
where $r$ is the exponential growth rate (and $\left[\psize(0)/\ssize(0)\right]^{-1} = \int_{-\infty}^{\infty} \exp(-r\tau') f_0(\tau') \dtau'$ since $b_0$ is a probability distribution).
For faster epidemics (higher $r$), the backward delay distribution will have a shorter mean.
Changes in the backward delay distribution are also by definition affected by the changes in the forward delay distribution (\eref{backward}).
Therefore, the backward delay distribution will generally differ from the forward delay distribution even if we are measuring time delays that are intrinsic to the life history of a disease (e.g., the incubation period).
These ideas apply to all epidemiological delay distributions and generalize the work by \cite{champredon2015intrinsic} who compared forward and backward generation-interval distributions to describe the realized generation intervals from the perspective of an infector and an infectee, respectively, as well as the work by \cite{britton2019estimation} who showed that \eref{backexp} holds for the backward generation-interval distributions.

\subsection{Realized serial interval distributions}

The serial interval is defined as the time between when an infector becomes symptomatic and when an infectee becomes symptomatic \citep{svensson2007note}.
Previous studies have often expressed serial intervals $\tau_{\rm s}$ in the form (\fref{diagram}A):
\begin{equation}
\tau_{\rm s} = (\gtime + \tau_{\rm i2}) - \tau_{\rm i1}
\end{equation}
where $\tau_{\rm i1}$ and $\tau_{\rm i2}$ represent incubation periods of an infector and
an infectee, respectively, and $\gtime$ represents the generation
interval.  
These studies assume that $\tau_{\rm i1}$ and $\tau_{\rm i2}$ are drawn from the same distributions and concluded that the serial and generation intervals have the same mean \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence, britton2019estimation};
however, this conclusion is based on the implicit assumption that distributions of realized incubation periods, $\tau_{\rm i1}$ and $\tau_{\rm i2}$, as well as generation interval, $\gtime$, are intrinsic to individuals (and not dependent on epidemic dynamics at the population-level) ---
something that is true of forward but not backward distribution of incubation periods.
We refer to this definition as the intrinsic
serial interval (\fref{diagram}A).

@@

To correctly link the realized serial interval distribution to the renewal process
between cases based on symptom onset dates, we must use the forward
serial interval (i.e., use the perspective of a cohort of infectors
that share the same symptom onset time).  Given that an infector
became symptomatic at time $\psymp$, to calculate the forward serial
interval we first go backward in time to when the infector was
infected, and then forward in time to when the infectee was infected,
and then forward to when the infectee became symptomatic.
In \fref{diagram}B, we see that $\tau_{\rm i1}$ is drawn from
the backward incubation period distribution of the cohort
of infectors who became symptomatic at time $\psymp$; $\gtime$ is drawn from
the forward generation-interval distribution of the cohort of
infectors who became infected at time $\psymp - \tau_{\rm i1}$; and $\tau_{\rm i2}$
is drawn from the forward incubation period distribution of the cohort of
infectees who became infected at time $\psymp - \tau_{\rm i1} + \gtime$.
Likewise, we can define the backward serial interval distribution for
a cohort of infectees who became symptomatic at time $\ssymp$
(\fref{diagram}C).  This conceptual framework demonstrates
that the distributions of $\tau_{\rm i1}$, $\gtime$, and $\tau_{\rm i2}$ (and therefore
the distributions of realized serial intervals) depend on the
reference cohort, which is defined by an event type (primary or secondary),
temporal direction (forward or backward), and a particular reference time.

To calculate realized serial-interval distributions, we begin by modeling $\total(\psymp,\ssymp)$: the total amount of serial intervals that start (when infectors develop symptoms) at time $\psymp$ and end (when infectees develop symptoms) at time $\ssymp$.
The amount of serial intervals between time $\psymp$ and $\ssymp$,
given that the infectors became infected at time $\pinf\le\psymp$ and
the infectees became infected at time $\sinf\le\ssymp$, depends on the
amount of infection that occurs between time $\pinf$ and $\sinf$ as
well as the density of forward incubation periods between $\pinf$ and
$\psymp$ (realized incubation periods of infectors) and between
$\sinf$ and $\ssymp$ (realized incubation periods of infectees):
\begin{equation}
\underbrace{\Rc (\pinf)}_{\substack{\text{case} \\ \text{reproduction} \\ \text{number}}} 
\times 
\underbrace{i(\pinf)}_{\text{incidence}} 
\times 
\underbrace{h_{\pinf}(\psymp-\pinf, \sinf - \pinf)}_{\substack{\text{joint density of} \\ \text{the forward incubation} \\ \text{period } p-\pinf \text{ and the forward} \\ \text{generation interval } \sinf-\pinf\\ \text{(of infectors)}}}
\times
\underbrace{\idist_{\sinf}(\ssymp - \sinf)}_{\substack{\text{marginal density of} \\ \text{the forward incubation} \\ \text{period } \ssymp-\sinf \\ \text{(of infectees)}}},
\end{equation}
where the case reproduction number $\Rc (\pinf)$ is defined as the average number of secondary cases caused by a primary case infected at time $\pinf$ over the course of their infection \citep{fraser2007estimating}.
We describe the forward incubation periods and the forward generation intervals using a joint probability distribution because onset of symptoms and transmission potential jointly depend on the life history of a disease;
for example, if an infected individual can only transmit the disease after symptom onset, the forward generation interval will necessarily be longer than the forward incubation period.

@@

During the initial phase of an epidemic, the proportion susceptible remains approximately constant ($S(t) \approx S(0)$) and incidence of infection grows exponentially: $i(t)=i_0\exp(rt)$.
During this period, the intrinsic generation-interval distribution provides the correct link between the exponential growth rate $r$ and the initial reproduction number $\RR=\Ro S(0)$ based on the Euler-Lotka equation \citep{wallinga2007generation};
in this case, we focus on the estimates of the basic reproduction number \Ro (the value of \RR in a fully susceptible population, $S(t) \approx 1$):
\begin{equation}
\frac{1}{\Ro} = \int_0^\infty \exp(-r\tau) g(\tau) \dtau.
\label{eq:Rgen}
\end{equation}
Analogous to the intrinsic generation-interval distribution, 
forward serial-interval distributions, by definition, describe the renewal process of infection based on symptom onset dates.
Therefore, we expect the forward serial-interval distribution during the exponential growth phase --- which we refer to as the \emph{initial} forward serial-interval distribution $f_0$ --- to estimate the same value of \Ro for a given $r$ as the intrinsic generation-interval distribution (note, however, that the forward serial interval is not necessarily positive):
\begin{equation}
\frac{1}{\Ro} = \int_{-\infty}^\infty \exp(-r\tau) f_{0}(\tau) \mathrm{d} \tau,
\label{eq:Rforward}
\end{equation}
where the initial forward serial-interval distribution is given by:
\begin{equation}
f_{0}(\tau) \propto \int_{-\infty}^{0} \int_{\pinf}^{\tau} \exp(r \pinf) h(-\pinf, \sinf - \pinf) \idist(\tau - \sinf) \, \mathrm{d}\sinf\,\mathrm{d}\pinf.
\label{eq:initialSI}
\end{equation}
In Supplementary Materials, we provide a mathematical proof of this relationship.
We further compare this with the estimate of \Ro based on the intrinsic serial-interval distribution $q(\tau)$:
\begin{equation}
\frac{1}{\Rintrinsic} = \int_{-\infty}^\infty \exp(-r\tau) q(\tau) \mathrm{d} \tau,
\label{eq:Rintrinsic}
\end{equation}
where the intrinsic serial-interval distribution $q(\tau)$ does not depend on epidemic dynamics:
\begin{equation}
q(\tau) \propto \int_{-\infty}^{0} \int_{\pinf}^{\tau} h(-\pinf, \sinf - \pinf) \idist(\tau - \sinf) \, \mathrm{d}\sinf\,\mathrm{d}\pinf.
\label{eq:intrinsicSI}
\end{equation}
Rather than numerically integrating over closed forms of $g$, $f_0$, and $q$ to estimate $\Ro$, we use simulation-based approaches for simplicity (Supplementary Materials).

The initial forward serial-interval distribution depends on the exponential growth rate $r$.
For a fast-growing epidemic (high $r$), we expect the backward incubation periods to be short (\eref{backexp}), and therefore, the forward serial-interval distribution will generally have a larger mean than the intrinsic generation- and serial-interval distributions.
The Susceptible-Exposed-Infected-Recovered model, under the additional assumption that the incubation and exposed periods are equivalent (i.e. that onset of symptoms and infectiousness occur simultaneously), provides a special case where the forward serial- and generation-intervals follow the same distributions during the exponential growth phase because (i) infected individuals can only transmit after symptom onset and (ii) the time between symptom onset to infection is independent of the incubation period of an infector (see Supplementary Materials).

@@

\subsection{Realized serial interval distributions}

@@

The initial forward serial interval distribution captures the exponential growth phase of an epidemic.
Next, we characterize how forward and backward serial intervals vary over the course of an epidemic by running deterministic and stochastic simulations based on the renewal equations (see Supplementary Materials) using parameters in Table 1;
we further assume $\Ro=2.5$.
While the forward serial-interval distribution is our primary focus, understanding the differences between the forward and the backward distributions is important because the observed intervals during an ongoing epidemic are often the backward ones:
we typically identify infected individuals and ask when and by whom they were infected.
Similarly, when we are estimating the incubation period of an individual, we typically observe their symptom onset date and try to estimate when they were infected (e.g., \cite{backer2020incubation}).

@@

The forward serial interval distributions depend on distributions of three intervals
(\fref{diagram}): (i) the backward incubation period, (ii) the forward generation interval, and (iii) the forward incubation period.
For COVID-19 parameters (Table 1), both forward incubation period (\fref{epi}B) and generation-interval (\fref{epi}C) distributions remain roughly constant;
therefore, changes in the forward serial-interval distributions are predominantly driven by changes in the backward incubation period distributions, whose mean increases over time (\fref{epi}D).
In this case, the increase in the mean backward incubation period directly translates to the decrease in the mean forward serial interval.
% Since the increase in the mean backward period is driven by the decreasing incidence, which can still occur in the absence of significant depletion of the susceptible pool via intervention, decrease in the mean forward serial interval may be a robust feature in COVID-19 outbreaks.
In general, relative contributions of the three distributions depend on their shapes, correlations between intrinsic incubation periods and generation intervals, and overall epidemiological dynamics.

@@

\subsection{Observed serial interval distributions}

@@

Now, we turn to issues of estimating the reproduction number from the observed serial-interval data during on ongoing epidemic.
In order to have an unbiased estimate of the basic reproduction number, we need to use serial-interval distributions estimated based on primary cohorts (i.e., infectors who share the same symptom onset time) at the early stage of the epidemic (i.e., initial forward serial-interval distribution).
However, researchers typically use all available information to estimate epidemiological parameters;
in particular, \cite{thompson2019improved} recently suggested that up-to-date serial-interval data are necessary to accurately estimate the reproduction number.
We explore the consequences of neglecting changes in the realized serial interval distribution on estimates of the basic reproduction number.

@@

While our results support the use of serial interval distributions for calculating \RR, 
they also reveal gaps in current practices in incorporating serial-interval distributions into outbreak analyses.
\cite{thompson2019improved} recently emphasized the importance of using up-to-date serial-interval data for accurate estimation of time-varying reproduction numbers.
However, our study demonstrates that early serial-interval data is required to estimate the initial \RR during the exponential growth phase and that using up-to-date serial-interval data (without accounting for changes in the forward serial-interval distribution) can, in fact, underestimate the initial \RR.
Future studies should explore how neglecting changes in the forward serial-interval distribution can affect the estimates of \RR beyond the exponential growth phase and potentially re-assess existing estimates of \RR.

Many studies have already estimated the (time-varying) reproduction numbers for COVID-19 epidemics using aggregated serial-interval data (e.g., \cite{tempvar,du2020serial,pan2020jama,zhang2020evolving});
these studies should also re-assess whether they appropriately considered the \emph{forward} serial interval.
Given large geographical variation in the spread of the COVID-19, the forward serial interval distributions are also likely to vary across countries.
We suggest that modelers should aim to characterize spatiotemporal variation in forward serial-interval distributions and understand their changes over time.
These modeling approaches should be necessarily coupled with intensive
epidemiological investigation through contact tracing at the very
beginning of the outbreak in order to characterize the initial forward
serial-interval distribution, particularly for an epidemic of an
emerging disease such as COVID-19.

>>>>

\jd{It's a shame not to mention the idea that serial intervals are wider, and should therefore predict smaller \RR. That would also help with understanding the resolution. Or maybe we get to these ideas later? It helps with the paradox framing, but I guess you really don't like that.}
One source of confusion arises from an apparent discrepancy between the serial-interval and generation-interval viewpoints.
When the epidemic is growing exponentially, the spread of infection can be characterized as a \emph{renewal process} based on previous incidence of infection, the associated generation-interval distribution, and the average infectiousness of an infected individual.
It is well established that this renewal formulation allows us to link the exponential growth rate of an epidemic $r$ with its reproduction number \RR using the generation-interval distribution \citep{wallinga2007generation}.
By definition, the serial-interval distribution describes the renewal process between symptomatic cases based on their symptom onset dates;
since both renewal processes, based on generation- and serial-interval distributions, describe the same underlying system of exponential growth, both should provide the same correct link between \RR and the epidemic growth rate $r$.
In contexts where the distributions are expected to be different, current theory has no explanation for how these differing distributions could provide identical estimates of \RR from $r$. 
Some studies have further suggested that using serial intervals can give different estimates of \RR \citep{britton2019estimation, ganyani2020estimating}.

@@

\eref{backward} suggests that backward delay distributions change over time even if their corresponding forward delay distribution does not change.
Backward delay distributions depend on changes in the primary cohort size over time due to conditionality of observations:
Conditioning on individuals whose secondary events have occurred at the same time means that we tend to observe shorter (or longer) inter-event delays when cohort size has been increasing (decreasing) through time.
When incidence is growing exponentially, we can calculate the amount of bias exactly.
Assuming that the forward delay distribution remains constant during the exponential growth phase ($f_p(\tau) \approx f_0(\tau)$), we can substitute $\psize(t) = \psize(0) \exp(rt)$ in \eref{backward} to obtain:
\begin{align}
b_0(\tau) &= \left[\psize(0)/\ssize(0)\right] \exp(-r\tau) f_0(\tau)
\label{eq:backexp}
\end{align}
where $r$ is the exponential growth rate (and $\left[\psize(0)/\ssize(0)\right]^{-1} = \int_{-\infty}^{\infty} \exp(-r\tau') f_0(\tau') \dtau'$ since $b_0$ is a probability distribution).
Thus, for faster epidemics (higher $r$), the backward delay distribution will have a shorter mean.
In general, the backward delay distribution will differ from the forward delay distribution (unless the disease is at equilibrium), even if we are measuring time delays that are intrinsic to the life history of a disease (e.g., the incubation period).
These ideas apply to all epidemiological delay distributions and generalize the work by \cite{champredon2015intrinsic} who compared forward and backward generation-interval distributions to describe realized generation intervals from the perspective of an infector and an infectee, respectively, as well as the work by \cite{britton2019estimation} who showed that \eref{backexp} holds for the backward generation-interval distributions.

\subsection{Realized serial-interval distributions}

The serial interval is defined as the time between when an infector becomes symptomatic and when an infectee becomes symptomatic \citep{svensson2007note}.
Previous studies have often expressed serial intervals $\tau_{\rm s}$ in the form (\fref{diagram}A):
\begin{equation}
\tau_{\rm s} = (\gtime + \tau_{\rm i2}) - \tau_{\rm i1}
\label{eq:intser}
\end{equation}
where $\tau_{\rm i1}$ and $\tau_{\rm i2}$ represent incubation periods of an infector and
an infectee, respectively, and $\gtime$ represents the generation
interval of the infector.  
These studies assume that $\tau_{\rm i1}$ and $\tau_{\rm i2}$ are drawn from the same distributions and concluded that the serial and generation intervals have the same mean \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence, britton2019estimation};
however, this conclusion is based on the implicit assumption that distributions of realized incubation periods, $\tau_{\rm i1}$ and $\tau_{\rm i2}$, as well as generation interval, $\gtime$, are intrinsic to individuals (and not dependent on epidemic dynamics at the population-level) ---
something that is generally true of forward but not backward incubation-period distributions.
We refer to the definition \eref{intser} as the intrinsic
serial interval (\fref{diagram}A).

@@

To correctly link the realized serial-interval distribution to the renewal process
between cases based on symptom onset dates, we must use the forward
serial interval (i.e., use the perspective of a cohort of infectors
that share the same symptom onset time).  Given that an infector
became symptomatic at time $\psymp$, to calculate the forward serial
interval we first go backward in time to when the infector was
infected, and then forward in time to when the infectee was infected,
and then forward to when the infectee became symptomatic.
In \fref{diagram}B, we see that $\tau_{\rm i1}$ is drawn from
the backward incubation period distribution of the cohort
of infectors who became symptomatic at time $\psymp$; $\gtime$ is drawn from
the forward generation-interval distribution of the cohort of
infectors who became infected at time $\psymp - \tau_{\rm i1}$; and $\tau_{\rm i2}$
is drawn from the forward incubation period distribution of the cohort of
infectees who became infected at time $\psymp - \tau_{\rm i1} + \gtime$.
Likewise, we can define the backward serial-interval distribution for
a cohort of infectees who became symptomatic at time $\ssymp$
(\fref{diagram}C).  This conceptual framework demonstrates
that the distributions of $\tau_{\rm i1}$, $\gtime$, and $\tau_{\rm i2}$ (and therefore
the distributions of realized serial intervals) depend on the
reference cohort, which is defined by an event type (primary or secondary),
temporal direction (forward or backward), and a particular reference time.

To calculate realized serial-interval distributions, we begin by modeling $\total(\psymp,\ssymp)$: the total amount of serial intervals that start (when infectors develop symptoms) at time $\psymp$ and end (when infectees develop symptoms) at time $\ssymp$.
The amount of serial intervals between time $\psymp$ and $\ssymp$,
given that the infectors became infected at time $\pinf\le\psymp$ and
the infectees became infected at time $\sinf\le\ssymp$, depends on the
amount of infection that occurs between time $\pinf$ and $\sinf$ as
well as the density of forward incubation periods between $\pinf$ and
$\psymp$ (realized incubation periods of infectors) and between
$\sinf$ and $\ssymp$ (realized incubation periods of infectees):
\jd{I feel that $R_c$ should group with h (i.e., after i()), for comprehension.}
\begin{equation}
\underbrace{\Rc (\pinf)}_{\substack{\text{case} \\ \text{reproduction} \\ \text{number}}} 
\times 
\underbrace{i(\pinf)}_{\text{incidence}} 
\times 
\underbrace{h_{\pinf}(\psymp-\pinf, \sinf - \pinf)}_{\substack{\text{joint density of} \\ \text{the forward incubation} \\ \text{period } p-\pinf \text{ and the forward} \\ \text{generation interval } \sinf-\pinf\\ \text{(of infectors)}}}
\times
\underbrace{\idist_{\sinf}(\ssymp - \sinf)}_{\substack{\text{marginal density of} \\ \text{the forward incubation} \\ \text{period } \ssymp-\sinf \\ \text{(of infectees)}}},
\end{equation}
where the case reproduction number $\Rc (\pinf)$ is defined as the average number of secondary cases caused by a primary case infected at time $\pinf$ over the course of their infection \citep{fraser2007estimating}.
We describe the forward incubation periods and the forward generation intervals using a joint probability distribution because onset of symptoms and transmission potential jointly depend on the life history of a disease;
for example, if an infected individual can only transmit the disease after symptom onset, the forward generation interval will necessarily be longer than the forward incubation period.

@@

During the initial phase of an epidemic, the proportion susceptible remains approximately constant ($S(t) \approx S(0)$) and incidence of infection grows exponentially: $i(t) \approx i_0\exp(rt)$.
During this period, the intrinsic generation-interval distribution provides the correct link between the exponential growth rate $r$ and the initial reproduction number $\RR=\Ro S(0)$ based on the Euler-Lotka equation \citep{wallinga2007generation};
in this case, we focus on the estimates of the basic reproduction number \Ro (the value of \RR in a fully susceptible population, $S(t) \approx 1$):
\begin{equation}
\frac{1}{\Ro} = \int_0^\infty \exp(-r\tau) g(\tau) \dtau.
\label{eq:Rgen}
\end{equation}
Analogous to the intrinsic generation-interval distribution, 
forward serial-interval distributions describe the renewal process of infection based on symptom onset dates.
Therefore, we expect the forward serial-interval distribution during the exponential growth phase --- which we refer to as the \emph{initial} forward serial-interval distribution $f_0$ --- to estimate the same value of \Ro for a given $r$ as the intrinsic generation-interval distribution (note, however, that the forward serial interval is not necessarily positive):
\begin{equation}
\frac{1}{\Ro} = \int_{-\infty}^\infty \exp(-r\tau) f_{0}(\tau) \mathrm{d} \tau,
\label{eq:Rforward}
\end{equation}
where the initial forward serial-interval distribution is given by:
\begin{equation}
f_{0}(\tau) \propto \int_{-\infty}^{0} \int_{\pinf}^{\tau} \exp(r \pinf) h(-\pinf, \sinf - \pinf) \idist(\tau - \sinf) \, \mathrm{d}\sinf\,\mathrm{d}\pinf.
\label{eq:initialSI}
\end{equation}
In Supplementary Materials, we provide a mathematical proof of this relationship.
We further compare this with the estimate of \Ro based on the intrinsic serial-interval distribution $q(\tau)$:
\begin{equation}
\frac{1}{\Rintrinsic} = \int_{-\infty}^\infty \exp(-r\tau) q(\tau) \mathrm{d} \tau,
\label{eq:Rintrinsic}
\end{equation}
where the intrinsic serial-interval distribution $q(\tau)$ does not depend on epidemic dynamics:
\begin{equation}
q(\tau) \propto \int_{-\infty}^{0} \int_{\pinf}^{\tau} h(-\pinf, \sinf - \pinf) \idist(\tau - \sinf) \, \mathrm{d}\sinf\,\mathrm{d}\pinf.
\label{eq:intrinsicSI}
\end{equation}
Rather than numerically integrating over closed forms of $g$, $f_0$, and $q$ to estimate $\Ro$, we use simulation-based approaches for simplicity (Supplementary Materials). Our parameter values are shown in the Table.

The initial forward serial-interval distribution depends on the exponential growth rate $r$.
For a fast-growing epidemic (high $r$), we expect the backward incubation periods to be short (\eref{backexp}), and therefore, the forward serial-interval distribution will generally have a larger mean than the intrinsic generation- and serial-interval distributions.
The Susceptible-Exposed-Infected-Recovered model, under the additional assumption that the incubation and exposed periods are equivalent (i.e. that onset of symptoms and infectiousness occur simultaneously), provides a special case where the forward serial- and generation-intervals follow the same distributions during the exponential growth phase because (i) infected individuals can only transmit after symptom onset and (ii) the time between symptom onset and infection is independent of the incubation period of an infector (see Supplementary Materials).

@@

\subsection{Realized serial-interval distributions}

@@

The initial forward serial-interval distribution captures the exponential growth phase of an epidemic.
Next, we characterize how forward and backward serial intervals vary over the course of an epidemic by running deterministic and stochastic simulations based on the renewal equations (see Supplementary Materials) using parameters in Table 1;
we further assume $\Ro=2.5$.
While the forward serial-interval distribution is our primary focus, understanding the differences between the forward and the backward distributions is important because the observed intervals during an ongoing epidemic are often the backward ones:
we typically identify infected individuals and ask when and by whom they were infected.
Similarly, when we are estimating the incubation period of an individual, we typically observe their symptom onset date and try to estimate when they were infected (e.g., \cite{backer2020incubation}).

@@

The forward serial-interval distributions depend on distributions of three intervals
(\fref{diagram}): (i) the backward incubation period, (ii) the forward generation interval, and (iii) the forward incubation period.
For COVID-19 parameters (Table 1), both forward incubation period (\fref{epi}B) and generation-interval (\fref{epi}C) distributions remain roughly constant;
therefore, changes in the forward serial-interval distributions are predominantly driven by changes in the backward incubation period distributions, whose mean increases over time (\fref{epi}D).
In this case, the increase in the mean backward incubation period directly translates to the decrease in the mean forward serial interval.
% Since the increase in the mean backward period is driven by the decreasing incidence, which can still occur in the absence of significant depletion of the susceptible pool via intervention, decrease in the mean forward serial interval may be a robust feature in COVID-19 outbreaks.
In general, relative contributions of the three distributions depend on their shapes, correlations between intrinsic incubation periods and generation intervals, and overall epidemiological dynamics.

@@

\subsection{Observed serial-interval distributions}

@@

Now, we turn to issues of estimating the reproduction number from the observed serial-interval data during on ongoing epidemic.
In order to have an unbiased estimate of the basic reproduction number, we need to use serial-interval distributions estimated based on primary cohorts (i.e., infectors who share the same symptom onset time) at the early stage of the epidemic (i.e., initial forward serial-interval distribution).
However, researchers typically use all available information to estimate epidemiological parameters;
in particular, \cite{thompson2019improved} recently suggested that up-to-date serial-interval data are necessary to accurately estimate the reproduction number.
We explore the consequences of neglecting changes in the realized serial-interval distribution on estimates of the basic reproduction number.

@@

While our results support the use of serial-interval distributions for calculating \RR, 
they also reveal gaps in current practices in incorporating serial-interval distributions into outbreak analyses.
\cite{thompson2019improved} recently emphasized the importance of using up-to-date serial-interval data for accurate estimation of time-varying reproduction numbers.
However, our study demonstrates that early serial-interval data is required to estimate the initial \RR during the exponential growth phase and that using up-to-date serial-interval data (without accounting for changes in the forward serial-interval distribution) can, in fact, underestimate the initial \RR.
Future studies should explore how neglecting changes in the forward serial-interval distribution can affect the estimates of \RR beyond the exponential growth phase and potentially re-assess existing estimates of \RR.

Many studies have already estimated the (time-varying) reproduction numbers for COVID-19 epidemics using aggregated serial-interval data (e.g., \cite{tempvar,du2020serial,pan2020jama,zhang2020evolving});
these studies should also re-assess whether they appropriately considered the \emph{forward} serial interval.
Given large geographical variation in the spread of the COVID-19, the forward serial-interval distributions are also likely to vary across countries.
We suggest that modelers should aim to characterize spatiotemporal variation in forward serial-interval distributions and understand their changes over time.
These modeling approaches should be necessarily coupled with intensive
epidemiological investigation through contact tracing at the very
beginning of the outbreak in order to characterize the initial forward
serial-interval distribution, particularly for an epidemic of an
emerging disease such as COVID-19.
